---
title: "Final_project"
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
date: "2025-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the dataset relavent to Experiment 1:

```{r}
library(dplyr)
library(tidyverse)
library(broom.mixed)
library(lme4)
library(ggeffects)
roettgerEtAlData <- read.csv("C:/Users/17390/Desktop/roettgerEtAlData.csv")
german <- roettgerEtAlData %>% 
  filter(subject >= 1 & subject <= 16)
```

### Build the Models
md1: model that examines the effect of voicing and other predictors on vowel duration (main effect model)
md1_interaction: model that examines the interaction of voicing with other predictors
md1_no_voicing: same as md1 but excludes voicing as a predictor

(1+ voicing|subject) + (1+ voicing|item_pair): random slopes and intercepts for subject and item_pair

```{r}
md1 = lmer(vowel_dur ~ voicing +	accent_type + prosodic_boundary +	
                place + vowel + norming_voiceless_count +
                (1+ voicing|subject) + (1+ voicing|item_pair),
              data=german)
```

### Summary of the main effect model:

```{r}
md1.tidy <- tidy(md1, conf.int = T)
md1.tidy
```
vowels estimated to be 8.7 ms longer before devoiced stops rather than to voiceless stops (SE = 2.03 ms)

### Plotting of the model
### plotting the predicted values
```{r}
mymodel <- predict_response(md1, terms = 'voicing', type = 'random')
plot(mymodel)
```
### plotting of the 95% CI
```{r}
md1.tidy %>% 
  ggplot(aes(x= term, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high),
                width = 0.5) +
  geom_hline(yintercept = 0, linetype = 2) +
  coord_flip()+
  labs(y = "Vowel Duration (ms)", x = "Predictors", title = "Plotting of 95% CI")
```

### Test 1: likelihood ratio test to compare md1 and md1_interaction
H0: The interaction between Voicing and other predictors does not have a significant effect on vowel duration
```{r}
md1_interaction = lmer(vowel_dur ~ voicing * (accent_type + prosodic_boundary +	
                       place + vowel + norming_voiceless_count) +
                       (1+ voicing|subject) + (1+ voicing|item_pair),
                     data=german)
anova(md1,md1_interaction,test="Chisq")
```
there were no interactions between VOICING and any of the control variables (χ2 (10) = 9.53, p = 0.48), H0 is accepted


### Test 2: likelihood ratio test to compare md1 and md1_no_voicing
Ho: Voicing does not have a significant effect on vowel duration
```{r}
md1_no_voicing = lmer(vowel_dur ~ accent_type + prosodic_boundary +
                 place + vowel + norming_voiceless_count + 		
                 (1+voicing|subject) + (1+voicing|item_pair),
               data=german)
anova(md1, md1_no_voicing, test = 'Chisq')
```
Voicing had a significant effect on vowel duration (χ2 (1) = 13.76, p<0.0003), H0 is rejected

### subset analyses:  exclude all not_usable responses
```{r}
german_filtered <- german %>% 
  filter(usable == "usable")
1 - nrow(german_filtered) / nrow (german)
```
removed 7.73% datapoints

### repeat the tests
```{r}
md2 = lmer(vowel_dur ~ voicing +	accent_type + prosodic_boundary +	
                place + vowel + norming_voiceless_count +
                (1+ voicing|subject) + (1+ voicing|item_pair),
              data=german_filtered)

md2_interaction = lmer(vowel_dur ~ voicing * (accent_type + prosodic_boundary +	
                       place + vowel + norming_voiceless_count +
                       (1+ voicing|subject) + (1+ voicing|item_pair)),
                     data=german_filtered)

md2_no_voicing = lmer(vowel_dur ~ accent_type + prosodic_boundary +
                 place + vowel + norming_voiceless_count + 		
                 (1+voicing|subject) + (1+voicing|item_pair),
               data=german_filtered)
```


```{r}
md1.tidy <- tidy(md1)
md1.tidy
```
vowels are 8.7 ms longer (SE = 2.03 ms) before devoiced stops.

```{r}
anova(md2,md2_interaction,test="Chisq")
```
After removing, there were still no interactions between VOICING and any of the control variables (χ2 (10) = 10.77, p = 0.376)

```{r}
anova(md2, md2_no_voicing, test = 'Chisq')
```
After removing, voicing still didn't have a significant effect on vowel duration (χ2 (1) = 13.37, p<0.0003)

```{r}
box_subject <- ggplot(german_filtered, aes(x=vowel_dur, y=subject)) + geom_point(color="darkred") + 
  ggtitle("Vowel Duration According to Subject") + xlab("Vowel Duration") + ylab("Subject") + 
  coord_flip()
```

```{r}
box_items <- ggplot(german_filtered, aes(x=vowel_dur, y=item_pair)) + geom_point(color="darkblue") + 
  ggtitle("Vowel Duration According to Items") + xlab("Vowel Duration") + ylab("Items") +
  coord_flip()
```

